Progress Report on One Simulated Dataset

### 2023-11-13 Docs/ProgressReportSim.Rmd
update figure

```{r, cache=TRUE}
# fit model on the training set
train_bin_lst <- split(train_df, f = train_df$bin)
# length(train_bin_lst)
# lapply(train_bin_lst, dim)

@@

# project mean functions 
# lm_mod <- lm(fpca_mod$mu ~ B-1)
# mu_new <- Bnew %*% coef(lm_mod)
```

@@

# subject 43, 79, 57, 148 have the most different prediction
df_pred %>% filter(id == id_diff$id[1]) %>% 
  select(t, starts_with("pred")) %>%
  pivot_longer(starts_with("pred")) %>%
  left_join(df_pred3 %>% filter(id == id_diff$id[1]) %>%
              select(t, starts_with("pred")) %>%
              pivot_longer(starts_with("pred")),
            by = c("t", "name")) %>%
  ggplot()+
  geom_point(aes(x=value.x, y=value.y), na.rm = T)+
  facet_wrap(~name)+
  labs(title = paste0("Subject ", id_diff$id[1]), x="Unscaled", y="Scaled")
```

@@

#### use Unscaled params ####
MyData_unscaled <- list(K=K, 
               X=efunctions_new[1:max_rid, ], 
               mon.names=mon.names,
               parm.names=parm.names, 
               pos.xi=pos.xi, 
               y=df_it$Y, 
               tao=diag(new_lambda), f0=new_mu[1:max_rid])
Fit_unscaled <- LaplaceApproximation(Model, parm = rep(0, K), Data=MyData,
                                     Method = "BFGS", Iterations = 1000,
                            CovEst = "Identity")

#### Scaled params ####
MyData_scaled <- list(K=K, 
               X=efunctions_scaled[1:max_rid, ], 
               mon.names=mon.names,
               parm.names=parm.names, 
               pos.xi=pos.xi, 
               y=df_it$Y, 
               tao=diag(scaled_lambda), f0=new_mu[1:max_rid])
Fit_scaled <- LaplaceApproximation(Model, parm = rep(0, K), 
                                   Data=MyData_scaled, Method = "BFGS", 
                                   Iterations = 1000,
                            CovEst = "Identity")

@@

```{r}
ggarrange(
Fit_unscaled$History %>% as.data.frame() %>%
  mutate(iter = 1:nrow(Fit$History)) %>%
  pivot_longer(starts_with("xi")) %>%
  left_join(true_xi_it, by = "name") %>%
  mutate(value = value/sqrt(n_bin)) %>%
  mutate_at(vars(value, true_xi), abs) %>%
  ggplot()+
  geom_point(aes(x=iter, y=value))+
  geom_line(aes(x=iter, y = true_xi, col = "red"))+
  facet_wrap(~name)+
  labs(title = "Unscaled parameter"),

@@

@@

>>>>

```{r}
# fit model on the training set
train_bin_lst <- split(train_df, f = train_df$bin)
# length(train_bin_lst)
# lapply(train_bin_lst, dim)

@@

@@

# subject 43, 79, 57, 148 have the most different prediction
df_pred %>% filter(id == id_diff$id[1]) %>% 
  select(t, starts_with("pred")) %>%
  pivot_longer(starts_with("pred")) %>%
  left_join(df_pred3 %>% filter(id == id_diff$id[1]) %>%
              select(t, starts_with("pred")) %>%
              pivot_longer(starts_with("pred")),
            by = c("t", "name")) %>%
  ggplot()+
  geom_point(aes(x=value.x, y=value.y), na.rm = T, size = 0.2)+
  facet_wrap(~name)+
  labs(title = paste0("Subject ", id_diff$id[1]), x="Unscaled", y="Scaled")
```

@@

# use unscaled params
MyData_unscaled <- list(K=K, 
               X=efunctions_new[1:max_rid, ], 
               mon.names=mon.names,
               parm.names=parm.names, 
               pos.xi=pos.xi, 
               y=df_it$Y, 
               tao=diag(new_lambda), f0=new_mu[1:max_rid])
Fit_unscaled <- LaplaceApproximation(Model, parm = rep(0, K), Data=MyData_unscaled,
                                     Method = "BFGS", Iterations = 1000,
                            CovEst = "Identity")

# use scaled params
MyData_scaled <- list(K=K, 
               X=efunctions_scaled[1:max_rid, ], 
               mon.names=mon.names,
               parm.names=parm.names, 
               pos.xi=pos.xi, 
               y=df_it$Y, 
               tao=diag(scaled_lambda), f0=new_mu[1:max_rid])
Fit_scaled <- LaplaceApproximation(Model, parm = rep(0, K), 
                                   Data=MyData_scaled, Method = "BFGS", 
                                   Iterations = 1000,
                            CovEst = "Identity")

@@

```{r}
ggarrange(
Fit_unscaled$History %>% as.data.frame() %>%
  mutate(iter = 1:nrow(Fit_unscaled$History)) %>%
  pivot_longer(starts_with("xi")) %>%
  left_join(true_xi_it, by = "name") %>%
  mutate(value = value/sqrt(n_bin)) %>%
  mutate_at(vars(value, true_xi), abs) %>%
  ggplot()+
  geom_point(aes(x=iter, y=value))+
  geom_line(aes(x=iter, y = true_xi, col = "red"))+
  facet_wrap(~name)+
  labs(title = "Unscaled parameter"),

@@

```{r converge_strata, eval=FALSE}
# Look at prediction results stratified by convergence status
converge_vec
converge_vec2
```

```{r, eval=FALSE}
# calcualte ISE
err1_conv <- df_pred %>%
  mutate(err1 = (pred0.2-eta_i)^2,
         err2 = (pred0.4-eta_i)^2,
         err3 = (pred0.6-eta_i)^2,
         err4 = (pred0.8-eta_i)^2) %>%
  select(id, t, starts_with("err"))

err1_conv$window = cut(err1_conv$t,
                       breaks = seq(0, 1, by = 0.2), include.lowest = T)
# table(err1$window)
err1_conv <- split(err1_conv, f = err1_conv$window)
err1_conv[[1]]

tb1 <- lapply(err1, function(x){
  x %>%  group_by(id) %>% 
    summarise_at(vars(err1, err2, err3, err4), sum) 
}) %>% bind_rows(.id = "Window")
```

@@

- Well, after increasing the scale of the true eigenvalues, the convergence rate did get closer between scaled and unscaled parameter. Also reversed as I expected.
- But on second though, A more effecient way is to calcualted ISE and AUC stratified by convergence status
